---
title: "Assignment 6"
author: "Megan Marziali"
date: "Feb 17, 2021"
output:
  github_document: default
---

## Assignment Set-Up

```{r message = FALSE}
library(tidyverse)
library(NHANES)
library(Amelia)
library(caret)

data(NHANES)
```

## Problem 1: Import and Restrict Data

```{r message = FALSE}
nhanes = NHANES %>% 
  janitor::clean_names() %>% 
  select(
    age, race1, education, hh_income, weight, height, 
    pulse, diabetes, bmi, phys_active, smoke100
  )
```

The NHANES data has 10,000 observations. To investigate missingness, I used the mapping function.

```{r message = FALSE, warning = FALSE}
missmap(nhanes)
```

It seems that education, smoking and pulse have a large amount of missing observations. However, I opted to keep all variables and exclude missing observations.

```{r message = FALSE}
nhanes_restr = nhanes %>% na.omit()
```

With missing observations remove, the total number of observations in this dataset is 6356. I next checked the balance of the outcome observations within the dataset:

```{r message = FALSE}
summary(nhanes_restr$diabetes)
```

There are 5697 "no" responses, and 659 "yes" responses, for a prevalence of diabetes within this sample of 11.6%.

```{r message = FALSE, warning = FALSE}
train.indices = createDataPartition(y = nhanes_restr$diabetes,p = 0.7,list = FALSE)

training = nhanes_restr[train.indices,]
testing = nhanes_restr[-train.indices,]
```




